Drawing Math

### 2022-05-12 paper/Paper.tex
last prop

\subsection{Finding period $T$}
\subsubsection{Finding the period of $\frac{n}{d}$}
We will first aim to find period $T$ of the state $\pmb{s}$.
For some $(n, b, d, \theta) \in \pmb{s}_0$, by \cite{MathOverflowFracPeriod},
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d.
\end{equation}
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$.

@@

\subsection{Property 1: Restricted Monomials and Closure}
Define $A_d = \set{j \mid j \in [k] \; \text{and}\; \inclIndic{d}{j} = 1}$, in other words,
$A_d$ is the set of rational parameters which are included in determining the position along the $d$th dimension.
Also, for function $f: \Z_\commonBase^{|A_d|} \rightarrow \Z_\commonBase$ and $\{a_1, a_2, ..., a_{|A_d|} \} = A_d$, we will denote
$$
  f\left(\digSumPeriodicNoJ^{a_1}, \digSumPeriodicNoJ^{a_2}, ..., \digSumPeriodicNoJ^{a_{A_d}}\right)
  = f\left(\digSumPeriodicNoJ\right).
$$

@@

If we were to then assume that $(\digSumPeriodicNoJ^1, ..., \digSumPeriodicNoJ^k)$ 
is uniformly and randomly draw from $\Z_\commonBase^k$, we then know by the 
Schwartz-Zippel Lemma
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) = 0\right] < \frac{2^k}{\commonBase}.
$$
% TODO: cite,
So, this would leave us with
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0\right] > 1 - \frac{2^k}{\commonBase}.
$$
In particular, this means that the probability of closure would be at least
$$
1 - \frac{2^k}{\commonBase}.
$$

@@

\section{Open Questions and Future Work}
A whole host of questions naturally arise from this problem. We will proceed to list
the ones which immediately stand out. For a complete list of open problems, which have
been solved, and the associated (financial) reward to each question, see our website
::::TODO::::
\begin{enumerate}
  \item \textbf{Periodic Digit Sum in NP?}: As far as the authors are aware, given 
  any rational number, $\frac{n}{d}$ in a fixed base $b$, is finding the sum of the digits in
  one period modulo $b$ in $NP$? I.e. could one give a poly-size
  certificate to a poly-time verifier which verifies if a digital sum equals $\sigma \in \Z_b$?
  The size of the input is the number of bits required to describe $\frac{n}{d}$. The authors 
  feel that intuitively the answer should be yes. A certificate could potentially be related
  to the factors of the denominator and the period length. I.e.\ we know how to find
  the periodic digital sum of a prime in polynomial time (see \ref{remark:digsum-complexity}). So,
  could we somehow use this in combination with a prime factorization?

@@

  \item \textbf{Distribution }: Given a random rational number $\frac{n}{d}$
  where $n \sim \Z_q$ and $d \sim \Z_q$ for some fixed $q \in \N$, what is the distribution
  of $\sigma$, the periodic digital sum modulo base $b$? Specifically, is the distribution
  indistinguishable (TODO: cite meaning) from the uniform distribution of $\Z_b$? 
  See remark \ref{remark:digSumRandomness} for an application of this result.

@@

@@

\appendix
\section{Proving Property 1}
\label{appendix:prop1}
\include{PaperAppendix.tex}

@@

% TODO: put b\theta / 2pi in State UPDATE

>>>>

\subsection{Finding period $T$}
We now show how to find a period $T$ of state $\pmb{s}$.

\subsubsection{Finding the period of a fraction}
For some $(n, b, d, \theta) \in \pmb{s}_0$,
we have that the period of the decimal expansion of $\frac{n}{d}$
can be determined by finding the smallest $\rationalPeriod$ such that
\begin{equation}
\label{eq:nd-period}
b ^ {\rationalPeriod} \equiv 1 \mod d
\end{equation}
by \cite{MathOverflowFracPeriod}.
More generally though, any nontrivial $\rationalPeriod$ satisfying equation \ref{eq:nd-period}
will be a period of $\frac{n}{d}$. In other words, the decimal sequence specified 
by $\frac{n}{d}$ will repeat after every $\rationalPeriod$ steps for all digits 
$n > n_0$ for some fixed $n_0$.

@@

\subsection{Property 1: Restricted Multinomials and Closure}
\label{prop:1}
Define $A_d = \set{j \mid j \in [k] \; \text{and}\; \inclIndic{d}{j} = 1}$, in other words,
$A_d$ is the set of rational parameters which are included in determining the position along the $d$th dimension.
Also, for function $f: \Z_\commonBase^{|A_d|} \rightarrow \Z_\commonBase$ and $\{a_1, a_2, ..., a_{|A_d|} \} = A_d$, we will denote
$$
  f\left(\digSumPeriodicNoJ^{a_1}, \digSumPeriodicNoJ^{a_2}, ..., \digSumPeriodicNoJ^{a_{A_d}}\right)
  = f\left(\digSumPeriodicNoJ\right).
$$

@@

If we were to then assume that $(\digSumPeriodicNoJ^1, ..., \digSumPeriodicNoJ^k)$ 
is uniformly and randomly draw from $\Z_\commonBase^k$, we then know that
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) = 0\right] < \frac{2^k}{\commonBase}
$$
by the Schwartz-Zippel Lemma \cite{Shwartz} \cite{Zip}.
So, this would leave us with
$$
  \Pr\left[\prod_{f \in \mathcal{M}} f(\digSumPeriodicNoJ) \neq 0\right] > 1 - \frac{2^k}{\commonBase}.
$$
In particular, this means that the probability of closure would be at least
$$
1 - \frac{2^k}{\commonBase}.
$$

@@

\subsection{Property 2: Root of Restricted Multinomial and Closure}
First note that
$$
\mathcal{M} = \{f : f(\pmb{x}) = \pm x_1 \pm x_2 ... \pm x_{|A_d|}\} = 
  \left\{\sum_{j \in |A_d|} (-1)^{\beta_j}\digSumPeriodic : \beta \in \set{0, 1}^{|A_d|} \right\}.
$$
Let $f_\beta \in \mathcal{M}$ then equal $\sum_{j \in |A_d|} (-1)^{\beta_j}\digSumPeriodic$.

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the Turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let rational sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the Turtle closes. If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$. Then $\exp(I \cdot r_i)$ can be thought of as some point around the 
unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given by a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the Turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

% TODO: picture!!!!

Over one common period for a all rational sequences, 
$\rationalPeriodTotal$,

\section{Open Questions and Future Work}
A whole host of questions naturally arise from this problem. We will proceed to list
the ones which immediately stand out. For a complete list of open problems, which have
been solved, and the associated (financial) reward to each question, see our website
\begin{enumerate}
  \item \textbf{Periodic Digit Sum in NP?}: As far as the authors are aware, given 
  any rational number, $\frac{n}{d}$ in a fixed base $b$, is finding the sum of the digits in
  one period modulo $b$ in $NP$? I.e. could one give a poly-size
  certificate to a poly-time verifier which verifies if a digital sum equals $\sigma \in \Z_b$?
  The size of the input is the number of bits required to describe $\frac{n}{d}$. The authors 
  feel that intuitively the answer should be yes. A certificate could potentially be related
  to the factors of the denominator and the period length. I.e.\ we know how to find
  the periodic digital sum of a prime in polynomial time (see \ref{remark:digsum-complexity}). So,
  could we somehow use this in combination with a prime factorization?

@@

  \item \textbf{Distribution }: Given a random rational number $\frac{n}{d}$
  where $n \sim \Z_q$ and $d \sim \Z_q$ for some fixed $q \in \N$, what is the distribution
  of $\sigma$, the periodic digital sum modulo base $b$? Specifically, is the distribution
  indistinguishable \cite{barak_2097} from the uniform distribution of $\Z_b$? 
  See remark \ref{remark:digSumRandomness} for an application of this result.

@@

  \item \textbf{Closure in BQP?}: Given that closure is intimately tied with period finding,
  the authors wonder if this problem is in BQP (Bounded-error Quantum Polynomial Time). An algorithm
  or existence result suffices.

  \item \textbf{A different approach?}: Could a change in basis and/ or vector calculus be used (as a means of dealing 
  with the change of basis from spherical coordinates to cartesian) to simplify the problem analysis?
  The authors conjecture yes as all updates are in spherical coordinates but most of the paper's analysis
  is done in cartesian space. But, it is not immediately obvious to the authors
  how a change of basis could be applied.

@@

\newpage
\include{PaperAppendix.tex}

@@
